### 2020-11-17 docs/modeling_fies_supplement.tex

\documentclass{article}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[a4paper, margin=1.25in]{geometry}
\usepackage{breakcites}
\usepackage{subcaption}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{textgreek}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{lineno}
\usepackage[
  style=numeric,
  citestyle=numeric-comp,
  backend=biber,
  doi=true,
  natbib=true,
  sorting=none
]{biblatex}

\pagestyle{fancy}
\fancyhf{}
\lfoot{Supplemental Materials for \textit{A Global Picture of Hunger}}
\rfoot{\thepage}

\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0pt}

\addbibresource{library.bib}

\begin{document}

\begin{center}
\section*{Supplemental Info for \textit{A Global Picture of Hunger}}
\end{center}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{section}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}

\section{Overview of the FIES}
We use raw microdata released by the FAO from 75 countries from different world regions and income categories (See Fig. \ref{fig:fies_countries}.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/FIES_Countries.pdf}
  \caption{Countries used in our model, shown in blue.}
  \label{fig:fies_countries}
\end{figure}

The microdata includes information about the probability that an individual is over the threshold for moderate and severe food insecurity, calculated by the FAO on the basis of the responses to the 8 FIES questions (See List \ref{itm:fies}).  The FIES is calculated using a Rasch model, which was developed by the psychometrics literature.  The Rasch model assumes that each individual and their responses to the FIES questions can be placed on a one-dimensional scale of food insecurity, and that the log odds of a respondent answering affirmatively to one of the FIES questions is a linear function of the difference between the severity of the food insecurity experienced by the individual and the severity of the item put forward in the corresponding question.

The severity of each item and the respondent’s level of food insecurity can be estimated making use of maximum likelihood methods. The assumptions of the Rasch model hold up well for the data at hand, as evidenced by model fit diagnostics \cite{Cafiero2018}.

After separately estimating the Rasch model for each country, FAO develops a global reference scale for the severity of each of the eight questions by iteratively harmonizing the severity values in each country. These global reference points are calibrated against other surveys that ask questions similar to those in the FIES, such as the HFSSM and the ELCSA.

Based on the global reference scale, two thresholds are set based on questions 1 and 8: whether the respondent is worried they would not have enough food to eat because of a lack of resources, and whether there was a time when they went without eating for a whole day because of a lack of money or other resources.

\begin{enumerate}
	\item During the last 12 MONTHS, was there a time when you were worried you would not have enough food to eat because of a lack of money or other resources?
	\item Still thinking about the last 12 MONTHS, was there a time when you were unable to eat healthy and nutritious food because of a lack of money or other resources?
	\item Was there a time when you ate only a few kinds of foods because of a lack of money or other resources?
	\item Was there a time when you had to skip a meal because there was not enough money or other resources to get food?
	\item Still thinking about the last 12 MONTHS, was there a time when you ate less than you thought you should because of a lack of money or other resources?
	\item Was there a time when your household ran out of food because of a lack of money or other resources?
	\item Was there a time when you were hungry but did not eat because there was not enough money or other resources for food?
	\item During the last 12 MONTHS, was there a time when you went without eating for a whole day because of a lack of money or other resources?
  \label{itm:fies}
\end{enumerate}

\section{Predicting Covariates Into the Future}
\subsection{Annualized Rate of Change}
For many of our covariates, to extrapolate recent trends to the future, we draw on historically annualized rates of change, which are used to project the future change in the variable. We use this method to estimate future rates of stunting, wasting, and malaria mortality, as well as to model future subnational shares of GDP and population.

For each subnational area, $a$, the rate of change (ROC) is calculated between each pair of adjacent years, $y$, based on a value, $p$, such that $0 < p < 1$, available for each year.

\begin{equation}
  \text{ROC}_{a,y} = \text{logit} \left( \frac{p_{a,y}}{p_{a,y-1}} \right)
  \label{eqn:a}
\end{equation}

Each ROC is weighted so as to give proportionally more weight to recent years.  For a dataset with observations for 2010 to 2017, that would be:

\begin{equation}
  W_y = \frac{y - 2010}{\sum_{y=2010}^{2017} y - 2010}
  \label{eqn:b}
\end{equation}

The annualized rate of change (AROC) is calculated as:

\begin{equation}
  \text{AROC}_{a} = \text{logit} \left( \sum_{y=2010}^{2017} W_y \times \text{AROC}_{a, y} \right)
  \label{eqn:c}
\end{equation}

Finally, the projections, Proj, for each year up to 2030 are given by:

\begin{equation}
  \text{Proj}_{a,y} = \text{logit}^{-1} ( \text{logit} ( p_{a,2017} ) + AROC_{a} \times ( y - 2017 ) )
  \label{eqn:d}
\end{equation}

For cases where we model shares with a summing constraint, such as the regional shares of a variable measured at the national level, we rescale the shares to ensure that the constraint holds.

\pagebreak
\subsection{Population}
Although population is not a covariate in our random forest model, it is an important determinant of other covariates such as GDP per capita. Future estimates of population are also necessary to convert modeled future rates of food insecurity into total headcounts of food insecurity.

We project population by combining historical subnational data from the Subnational Development Database \citep{Smits2019} with national level population estimates for the 21st century \citep{KC2017}.  To account for within-country trends and changes in population distribution, we use the AROC method outlined in equations \ref{eqn:a} - \ref{eqn:d} to project each subnational areas share of national population totals. We then disaggregate the national totals given by KC et al. to estimate future subnational populations.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/population.png}
  \caption{Population}
\end{figure}

\pagebreak
\subsection{Urbanization Rates}

We draw our historical and future estimates of urbanization entirely from the spatially explicit scenarios of urban population by Jones and O'Neill \citep{Jones2016}, which provide a series of urbanization rate projections which are consistent with the Shared Socioeconomic Pathways.  As with our other covariates, we use projections consistent with SSP2, the middle-of-the-road scenario.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/urban_perc.png}
  \caption{Percentage of people living in urban areas.}
\end{figure}

\pagebreak
\subsection{Wasting}
To estimate the prevalence of wasting for each administrative area in our dataset, we use data from the Local Burden of Disease project.  For the years 2010 to 2017, we simply use the mean rate of wasting in each administrative area from the dataset.  Higher-income countries that are not included in the dataset are assumed to have zero wasting.  We then project wasting for the years 2018-2030 using the AROC method, which the Local Burden of Disease group similarly use to estimate wasting for the year 2025 \citep{Local2020}.

To account for the effects of the coronavirus pandemic on global rates of wasting, we assume that long-term trends in rates of wasting would hold steady, after they increase globally by 14.3\% in 2020, based on estimates published in \textit{The Lancet} \citep{headey2020impacts}.  We model this impact as uniform across all countries where wasting occurs.  We project the rate of wasting in each administrative area in 2021 as being the average of (a) the rate of change in 2020 accounting for the pandemic shock and (b) the predicted rate of change for 2022 obtained ignoring the pandemic.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/wasting.png}
  \caption{Prevalence of wasting.}
\end{figure}

\pagebreak
\subsection{Stunting}
We model stunting, sourced from the Local Burden of Disease dataset, using the same methodology employed for wasting. We also assume a 14.3\% increase in prevalence for the year 2020 \citep{headey2020impacts}, as is done for wasting, although stunting is a long-term consequence of malnutrition and is not likely to be observable in a population with the same lags as wasting would be. We use stunting in our modeling framework as a proxy for chronic conditions of hunger, which are probably exacerbated by the coronavirus pandemic.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/stunting.png}
  \caption{Prevalence of stunting.}
\end{figure}

\pagebreak
\subsection{GDP Per Capita}
We estimate global subnational GDP per capita using historic subnational estimates of Gross National Income (GNI) from the Subnational Human Development Database \cite{Smits2019} and forecasts of national GDP within the SSP framework from Dellink et al. \cite{Dellink2017}.  We first use data from the World Bank to estimate the mean country-level ratio of GNI to GDP and adjust the subnational estimates of GNI according to that country-specific ratio.  We then calculate each subnational area's share of country GDP, and then use the AROC method to estimate each subnational area's share of total country GDP.  Based on these trajectories in each subnational areas share of total country GDP, we estimate what each subnational area's share of GDP will be for each year from 2019 to 2030.  We divide that share of GDP by the population of the administrative area, which we had previously estimated.  Thus, national GDP totals are consistent with the projections of Dellink et al., but are disaggregated according to the historic patters of GDP share in the Subnational Human Development Database.

Finally, we used estimates of country-level changes in GDP growth for the years 2020 and 2021 to account for the impact of the coronavirus \citep{prospects2020}.  We then model GDP as resuming its previous growth trajectory based on previously-estimated growth rates from 2022-2030, but growing from an estimate for 2021 that includes the coronavirus shock.

We measure global subnational GDP per capita using historic subnational estimates of Gross National Income (GNI) from the Subnational Human Development Database \citep{Smits2019} and forecasts of national GDP within the SSP framework form Dellink et al. \citep{Dellink2017}. We source data on the mean country-level ratio of GNI to GDP from the World Bank and adjust the subnational estimates of GNI according to the ratio for each country in our sample. We calculate the share of national GDP for each subnational region, and use the AROC method to project each subnational area’s share of total country GDP up to 2030. National GDP totals are consistent with the projections of Dellink et al., but are disaggregated according to the historic patterns of subnational GDP shares sourced from the Subnational Human Development Database. We use nowcasts and forecasts of country-level changes in GDP growth for the years 2020 and 2021 to account for the impact of the coronavirus \citep{prospects2020}. After the shock created by the COVID pandemic, GDP is assumed to resuming its previous growth trajectory for the period 2022-2030.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/gdp_percap.png}
  \caption{GDP per capita.}
\end{figure}

\pagebreak
\subsection{Mean Years of Schooling}
We use data on global subnational mean years of schooling from the Subnational Human Development Database \cite{Smits2019} and forecasts of national mean years of schooling from KC et al. \cite{KC2017}.  Using the subnational data, we calculate the difference in years of schooling form the nation mean each subnational administrative area, and use this value to disaggregate the future projections to a subnational level, keeping these differences constant over the projection horizon.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/school_mean.png}
  \caption{Mean years of schooling.}
\end{figure}

\pagebreak
\subsection{Gini Coefficient}
We use data on historical levels and projections of national income inequality from Rao et al. \citep{Rao2019a}.  These estimates are not disaggregated to a subnational level.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/gini.png}
  \caption{Gini Coefficient}
\end{figure}

\pagebreak
\subsection{Poverty Headcount Index}
We used historical and projected data on the Poverty Headcount Index (using the standard threshold of \$1.90 a day) from the World Poverty Clock by the World Data Lab, which in turn uses Povcal data sourced from the World Bank \citep{Cuaresma2018}. These data are the version corresponding to the update in summer 2020 and include the expected impact of the COVID pandemic on global poverty dynamics. Using parameterized Lorenz curves in combination with mean income/consumption and population projections, poverty rates are projected.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/hci.png}
  \caption{Poverty Headcount Index}
\end{figure}

\pagebreak
\subsection{Water Scarcity}
We use data for the water scarcity index (WSI) developed by \citep{greve2018global}. The index represents the ratio of the decadal averages in water demand to water supply, based on population projections consistent with SSP2 and water availability projections from global hydrological and general circulation models.  Based on this index, we estimate the share of people living in areas with less than 1000m\textsuperscript{3} of water available per capita per year. The WSI has been used already to build the water scarcity clock.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/ws_share.png}
  \caption{Water Scarcity Share}
\end{figure}

\pagebreak
\subsection{Mean Annual Precipitation and Mean Temperature}
For mean annual precipitation and temperature, we use data for the years 2010 to 2019 from the TerraClimate dataset \citep{abatzoglou2018terraclimate}, aggregated to each subnational area.  For data for the years 2020-2030, we use projections from models from the Inter-Sectoral Model Intercomparison Project \citep{warszawski2014inter}.  Specifically, we use the mean of an ensemble of bias-corrected projections under representative concentration pathway (RCP) 6.0.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/precip.png}
  \caption{Mean Annual Precipitation}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/tave.png}
  \caption{Mean Temperature (Celsius)}
\end{figure}

\pagebreak
\subsection{Topographic Ruggedness}
As a measure of accessibility, our models include the mean topographic ruggedness of each subnational area as a covariate. We obtain the variable using a gridded dataset of elevation from the USGS \citep{USGS1996} and calculate the index at each grid cell using the methodology from Riley et al. \citep{Riely1999}. We then aggregate the values for all the grid cells within each administrative area. The variable is constant in the estimation and projection period.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/ruggedness.png}
  \caption{Topographic ruggedness}
\end{figure}

\pagebreak
\subsection{Malaria (\textit{P. falciparum}) Mortality Rate}
We used data from Weiss et al. to estimate the mortality (deaths per 100,000 population per year) due to Malaria \citep{Weiss2019}, and aggregate the information at the level of subnational administrative area. Projections of malaria mortality are based on the AROC method.

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/covars/mal_falciparum.png}
  \caption{Rate of mortality due to \textit{P. falciparum} Malaria}
\end{figure}

\pagebreak
\section{Description, Implementation and Validation of the Random Forest Regression Model}
The RF regression is a machine learning method based on the creation of a large number of decision trees. By taking an ensemble of decision tree models, random forests introduce more variance and balance out the bias that is common to methods based on single decision trees \citep{friedman2001elements}.  For each of the decision trees in a random forest model, observations are selected at random with replacement, a method known as bootstrap aggregating, or bagging, and features are also selected at random.

For the implementation we use the R-package \texttt{randomForestSRC} by Ishwaran and Kogalur \citep{ishwaran2019randomforestsrc}. Specifically, we apply the functions \texttt{tune.rfsrc()} to find the optimal combination of hyperparamters and the function \texttt{rfsrc()} to perform the RF regression models. After predicting rates of moderate-to-severe and severe food insecurity, we plot the modeled vs. observed rates and calculated the MSE and ${R}^2$ (See Fig. \ref{fig:rf_in-sample} and Fig. \ref{fig:rf_out-sample}) for both models. Additionally, we use the function \texttt{vimp()} to gain insights into the development of the error rates with increasing numbers of trees (See Fig. \ref{fig:rf_error}) and the importance of individual variables (See Fig. \ref{fig:rf_vimp}).

%STEP-BY-STEP: LOGIT TRANSFORM, TUNE, MODEL, PREDICT, VALIDATE
Because the outcome variables in our models, the rate of moderate or severe food insecurity, are a bounded outcome, we first used a logistic transformation to extend the range from $[0,1]$ to $(-\infty, \infty)$.  Then, after estimating the model, we used the inverse logit to convert the unbounded model predictions to a rate between 0 and 1.

To ensure that predicted rates are bounded by 0 and 1 we conduct a logit transformation of the dependent variable and inverse logit transformation of the predictions. 
%MAYBE STATE THIS SOMEWHERE ELSE, NOT SURE WHERE

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/model/error_rf.png}
  \caption{Error Rate of the RF regression model. Panel (A) shows the Moderate-to-Severe model and Panel (B) shows the Severe model.}
  \label{fig:rf_error}
\end{figure}

%HERE A FEW SENTENCES ON THE NUMBER OF TREES WITH LINK TO NEXT PARAGRAPH

In addition to the parameter that describes the number of trees to be created, two other important hyperparameters must be tuned. In the function \texttt{tune.rfsrc()} these parameters are called \texttt{mtry} and \texttt{nodesize}. \texttt{mtry} describes the number of variables randomly selected as candidates for splitting a node and \texttt{nodesize} describes the average number of observations in a leaf node. The optimal pair of hyperparamters is found by trying different combinations and choosing the one with the smallest out-of-bag (OOB) error.

%EXPLAIN THE OOB ERROR HERE?

The function \texttt{tune.rfsrc()} does exactly that and returns optimal values for \texttt{mtry} and \texttt{nodesize}. We applied this function on both, the moderate-to-severe and the severe model. For the moderate-to-severe model the optimal combination is \texttt{mtry} = 12 and \texttt{nodesize} = 1 and for the severe model we found \texttt{mtry} = 10 and \texttt{nodesize} = 2 to be optimal.

%STATE FUNCTION ARGUMENTS HERE?

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/model/in-sample_rf.png}
  \caption{In-Sample Fit of the RF regression model. Panel (A) shows the Moderate-to-Severe model and Panel (B) shows the Severe model.}
  \label{fig:rf_in-sample}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/model/out-sample_rf.png}
  \caption{Out-Of-Sample Fit of the RF regression model with 75\% training set and 25\% test set. Panel (A) shows the Moderate-to-Severe model and Panel (B) shows the Severe model.}
  \label{fig:rf_out-sample}
\end{figure}

%HERE EVERYTHING ON THE MODEL, NOT SURE HOW DETAILED
%For modeling the rates of moderate-to-severe and severe food insecurity we use the function \texttt{rfsrc()}. We specify

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/VIMP.pdf}
  \caption{Variable Importance of the RF regression model. Panel (A) shows the Moderate-to-Severe model and Panel (B) shows the Severe model.}
  \label{fig:rf_vimp}
\end{figure}

%https://kogalur.github.io/randomForestSRC/theory.html#section4.2
Lastly we asses the importance of individual variables. This is based on the Breiman-Culter permutation variable importance information \citep{breiman2001random}.  This involved comparing the prediction error on the OOB data to the prediction error of OOB cases where the x-variable $x$ is randomly permutated.  The importance of an individual variable is thus the mean difference between the perturbed and unperturbed error rate, across all trees. 
%MORE INFORMATION HERE
\end{document}

\printbibliography